---
title: "dashboard ideas/exploration"
format: html
editor_options: 
  chunk_output_type: console
---

# load libraries
```{r}
library(tidyverse)
library(here)
library(sf)
library(tmap)
library(stars)
library(terra)
```

# read in data
```{r}
historical <- read_csv(here("data", "historical_production_delivery.csv"))
metadata <- read_csv(here("data", "metadata.csv"))
dictionary <- read_csv(here("metadata", "historical_production_delivery_data_dictionary.csv"))
actual_shortage <- read_csv(here("data", "actual_water_shortage_level.csv"))
fiveyr_shortage <- read_csv(here("data", "five_year_water_shortage_outlook.csv"))
monthly_shortage <- read_csv(here("data", "monthly_water_shortage_outlook.csv"))
```

```{r}
boundary <- read_sf(here("data", "supplier_shape", "California_Drinking_Water_System_Area_Boundaries.shp"))
```

```{r}
tm_shape(boundary) + 
  tm_polygons()
```

## Historical dataset:

```{r}
historical_nona <- historical |> 
  filter(quantity_acre_feet >= 0) |> 
  filter(water_produced_or_delivered == "water produced")

historical_small = historical_nona |> 
  filter(pwsid == "CA3610001")

historical_small2 = historical_nona |> 
  filter(pwsid == "CA2710001")
```

```{r}
ggplot(historical_small, aes(x = start_date, y = quantity_acre_feet, color = pwsid)) +
  geom_line()
```

```{r}
ggplot(historical_small2, aes(x = start_date, y = quantity_acre_feet, color = pwsid)) +
  geom_line()
```

```{r}

historical_compare_del <- historical |> 
  filter(quantity_acre_feet >= 0) |> 
  filter(water_produced_or_delivered == "water delivered") |> 
  rename(delivery_quantity = quantity_acre_feet) |> 
  select(water_system_name, delivery_quantity, pwsid, start_date)

historical_compare_prod <- historical |> 
  filter(quantity_acre_feet >= 0) |> 
  filter(water_produced_or_delivered == "water produced") |> 
  rename(production_quantity = quantity_acre_feet) |> 
  select(-water_produced_or_delivered) 


historical_small2 = historical_compare_del |> 
  filter(pwsid == "CA3610001")

historical_small3 = historical_compare_prod |> 
  filter(pwsid == "CA2710001")

```


```{r}
ggplot(historical_small2, aes(x = start_date, y = delivery_quantity)) +
  geom_line()

ggplot(historical_small3, aes(x = start_date, y = production_quantity)) +
  geom_line()


```

possible variables of interest:

we might want to visualize date (start_date?) with production or delivery quantity

for example: in the dashboard, if we select a specific pwsid, we could show the graph of production quantity over time. 

Maybe also be able to select water type and visualize average production/delivery over time

example of what the dashboard graph could look like:
```{r}
ggplot(historical_small3, aes(x = start_date, y = production_quantity)) +
  geom_line(color = "navy") +
  labs(
    title = "City of Adelanto Water System's Historical Production ",
    x = "Date (2013-2022)",
    y = "Production Amount (acre-feet)"
  ) +
  theme_bw()+
  scale_x_date(expand = c(0, NA)
  ) +
  scale_y_continuous(expand = c(0, NA))
```

## Actual shortage dataset

```{r}
# join with historical?

actual_shortage_2022 <- actual_shortage |> 
  mutate(year = year(start_date)) |> 
  filter(year == 2022) 

ggplot(actual_shortage_2022, aes(x = start_date, y = state_standard_shortage_level)) +
  geom_col()

```


